% -*- mode: LaTeX; TeX-PDF-mode: t; -*-
\input{@resources/tex-add-search-paths}  % allow latex to find custom stuff
\input{./.econtexRoot}\documentclass[\econtexRoot/BufferStockTheory]{subfiles}

\newcommand{\subname}{Conclusions}
\input{\econtexRoot/@local/dir-paths}
\usepackage{econark-ifsubfile}
\usepackage{econark-xrsetup}
\externaldocument{\LaTeXGenerated/BufferStockTheory}

\xrsetup{\econtexRoot/\texname}

\begin{document}

\newpage
\hypertarget{Conclusions}{}
\section{Conclusions}

Numerical solutions to optimal consumption problems, in both life cycle and infinite-horizon contexts, have become standard tools since the first reasonably realistic models were constructed in the late 1980s.
One contribution of this paper is to show that finite-horizon (`life cycle') versions of the simplest such models, with assumptions about income shocks (transitory and permanent) dating back to~\cite{friedmanATheory} and standard specifications of preferences --- and without plausible (but computationally and mathematically inconvenient) complications like liquidity constraints --- have attractive properties (like continuous differentiability of the consumption function, and analytical limiting MPC's as resources approach their minimum and maximum possible values).%, and that (more widely used) models with liquidity constraints can be viewed as a particular limiting case of this simpler model.

The main focus of the paper, though, is on the limiting solution of the finite-horizon model as the time horizon approaches infinity.
This simple model has other appealing features: A \hyperlink{FVAC}{`Finite Value of Autarky'} condition guarantees convergence of the consumption function, under the mild additional requirement of a \hyperlink{WRIC}{`Weak Return Impatience Condition'} that will never bind for plausible parameterizations, but provides intuition for the bridge between this model and models with explicit liquidity constraints.
The paper also provides a roadmap for the model's relationships to the perfect foresight model without and with constraints.
The constrained perfect foresight model provides an upper bound to the consumption function (and value function) for the model with uncertainty, which explains why the conditions for the model to have a non-degenerate solution closely parallel those required for the perfect foresight constrained model to have a non-degenerate solution.

The main use of infinite-horizon versions of such models is in heterogeneous-agent macroeconomics.
The paper articulates intuitive `Growth Impatience Conditions' under which populations of such agents, with Blanchardian (tighter) or Modiglianian (looser) mortality will exhibit balanced growth.
Finally, the paper provides the analytical basis for many results about buffer-stock saving models that are so well understood that even without analytical foundations researchers uncontroversially use them as explanations of real-world phenomena like the cross-sectional pattern of consumption dynamics in the Great Recession.

%% % The paper's results are all easily reproducible \href{https://econ-ark.org/_materials/BufferStockTheory?launch}{interactively on the web} or \href{https://github.com/econ-ark/BufferStockTheory}{on any standard computer system}.  Such reproducibility reflects the paper's use of the open-source {\ARKurl} toolkit, which is used to generate all of the quantitative results of the paper, and which integrally incorporates all of the analytical insights of the paper.

%% % The Dummy equation below sems to be needed to get the equation numbering in the appendix
%% % reliably to start at the next number after the last actual equation number used in the paper

%% \clearpage\vfill\eject
%% \begin{equation*}
%%   \label{eq:Dummy}
%% \end{equation*}

%% \notinsubfile{\bibliography{
%% \texname, % subfile inherits texname from preamble of parent
%% \econtexBib % Default bib database is in Resources/LaTeXInputs
%% }}

%% \makeatletter
%% \@ifundefined{subfiles@main}{%
%%   % Standalone mode
%%   \econarkmultibib{BufferStockTheory}
%% }{%
%%   % Included in main document
%%   \renewcommand{\bibdata}[1]{\@input@{\jobname.bbl}}
%%   \econarkmultibib{BufferStockTheory}
%% }
%% \makeatother

% documentclass is {article} or something else when main is compiling
% \ifSubfilesClassLoaded{
%     % Content you want to compile only when standalone.
%     \providecommand{\bibfiles}{}\renewcommand{\bibfiles}{\econarkcollectbibs{\econtexRoot/\texname}}\bibliography{\bibfiles}
% }{} % end ifSubfilesClassLoaded

\end{document}\endinput
